---
title: "Problem Set 12: NYC Complaints Data Analysis"
author: "<Judah Altman>"
date: "`r Sys.Date()`"
format: html
---

```{r}
#| label = "load-libraries-data",
#| warning = FALSE,
#| message = FALSE
library(tidyverse)
library(here)
library(socviz)

## Mapping
library(sf)
library(ggpattern)
library(patchwork)

## Census
## Make sure you have a census API key! You only need one.
## https://api.census.gov/data/key_signup.html
library(tidycensus)
options(tigris_use_cache = TRUE)
options(scipen = 999)

## Install the data package:
## remotes::install_github("kjhealy/nycomplaints")
library(nycomplaints)
```

# The dataset

The dataset is documented [on its website](https://kjhealy.github.io/nycomplaints/). The source data is from [NYC Open Data](https://data.cityofnewyork.us/City-Government/NYC-Council-Constituent-Services/b9km-gdpy/about_data).

Read the documentation before working with the data.

# Explore the data

## The complaints data:

```{r}
nycomplaints
```

## Two useful tables

```{r}
nyc_zips
```

```{r}
census_vars
```

# Things to do

## 1. Briefly Describe the Dataset

-   Summarize the `nycomplaints` tibble in a way you think is informative to you.

```{r}
head(nycomplaints)
ncol(nycomplaints)
colnames(nycomplaints)
nrow(nycomplaints)
length(unique(nycomplaints$unique_key))
```

This is a tibble of constituent complaints registered by NYC Council district offices. It contains 341,299 rows and 11 columns, where each row (for the most part, to be investigated below) is a complaint. By columns, the unique key column identifies each complaint, the account refers to the council district that the complaint was filed in, the open and close date refer to the respective dates that the complaint was filed and when the case was closed, the complaint type categorizes the type of complaint, the descriptor gives a specific category within the broader type, and the remaining variables give the zip, borough, city, council district and community board of the constituent (note that the council district listed here might be different from the council district listed under account, as the account district refers to which district the complaint was filed in, while the council district refers to the council district that the constituent that filed the complaint lives in).

## 2 Look at each column/variable in a little more detail

-   Explore the basic structure of the data, by e.g. writing code to summarize the variables or investigate aspects of them that seem interesting. You can also create some initial rough exploratory plots if you like, in addition to tables or other summaries. As you go, explain here what it is you are doing and why.

```{r}
#overall data structure - checking the number of rows and columns and the types of data in the tibble
head(nycomplaints)
ncol(nycomplaints)
colnames(nycomplaints)
nrow(nycomplaints)

nycomplaints %>% 
  map(class)

#unique ID - checking for duplicate rows
length(unique(nycomplaints$unique_key))
nrow(nycomplaints)
duplicate_keys <- nycomplaints %>% 
  filter(duplicated(nycomplaints$unique_key)) %>% 
  select(unique_key) %>% 
  pull() %>% 
  str_c(collapse = "")

#checking if they are actually duplicates (they are)
duplicate_nycomplaints <- nycomplaints %>% 
  filter(str_detect(duplicate_keys, unique_key))
View(duplicate_nycomplaints)

#account - what are the different council districts - there are 51
nycomplaints %>% 
  group_by(account) %>% 
  summarize() %>% 
  print(n = Inf)

#check council district here (should be the same as account), but ends up being different than account - there are 157 council districts, perhaps reflects a wider set of council districts as compared to council districts that file complaints, as the council districts here, except for a few coding errors, go all the way to 103 - do need to clean up some of the council names
nycomplaints %>% 
  group_by(council_dist) %>% 
  summarize() %>% 
  arrange() %>% 
  print(n = Inf)

#earliest and latest date of opening - data set runs from January 1st 2015 to January 9 2025 for opened files
nycomplaints %>% 
  select(opendate) %>% 
  filter(opendate == min(opendate) | opendate == max(opendate))

#earliest and latest date of closing - data set runs from January 2nd 2015 to January 9 2025, there are also some NA values for closing likely corresponding to non-closed cases
nycomplaints %>% 
  select(closedate) %>% 
  drop_na() %>% 
  filter(closedate == min(closedate) | closedate == max(closedate))

#looking at complaint types - very uneven coding of complaints, with many that makes sense, but also many with spelling errors
nycomplaints %>% 
  group_by(complaint_type) %>% 
  summarize() %>% 
  print(n = Inf)

#looking at descriptors - too many to summarize so split by complaint type and see how many descriptors are in each complaint type
nycomplaints %>% 
  group_by(descriptor) %>% 
  summarise()
  
descriptors_by_complaint_type <- nycomplaints %>% 
  group_by(complaint_type) %>% 
  reframe(unique(descriptor))

#checking for zip - many weird zips that aren't actually zips
nycomplaints %>% 
  group_by(zip) %>% 
  summarise() %>% 
  print(n = 100)

#borough - normal except for presence of NAs
nycomplaints %>% 
  group_by(borough) %>% 
  summarise()

#city - very odd, innacurate and messy, over 1400 different inputs
nycomplaints %>% 
  group_by(city) %>% 
  summarise()

#checking community board, mostly matches boroughs subset by board number but with some numbers that don't have a borough
nycomplaints %>% 
  group_by(community_board) %>% 
  summarise() %>% 
  arrange() %>% 
  print(n = Inf)

#tidying
#keeping only distinct ID keys (removing duplicates), filtering out non-existent zips,  changing council districts to match account district patterns, merging the table with nyc_zips to clean up borough and city names
nycomplaints_clean <- nycomplaints %>% 
  distinct(unique_key, .keep_all = TRUE) %>% 
  filter(zip %in% nyc_zips$zip) %>% 
  mutate(zip = as.numeric(zip)) %>% 
  mutate(council_dist = str_replace_all(council_dist, "NYCC0..", "NYCC.."))

#merging nyc_zips and nycomplaints_clean to clean up the borough and city names
nycomplaints_clean <- nycomplaints_clean %>% 
  left_join(nyc_zips, by = "zip") %>% 
  select(-borough.x, -city.x)
```

## 3. Make some plots or maps

-   Following on from the tabular summarize, make some plots that you think show something interesting about the data. Make sure they look presentable and effective to you. Make a note about why you're doing each one. If you like, merge in data from elsewhere (e.g. from Census variables).

```{r}
#loading in census data already generated in the import_tidy file
nyzip_census_wide <- readRDS(here("nyzip_census_wide.rds"))
nyzip_census <- readRDS(here("nyzip_census.rds"))

#merging the two tibbles by zip
nycomplaints_spatial <- nycomplaints_clean %>% 
  left_join(nyzip_census_wide, by = "zip")

#creating a zip and geometry tibble
zip_geometry <- nycomplaints_spatial %>% 
  group_by(geometry, zip) %>% 
  summarize()

#pulling out information about complaints by zip
complaints_by_zip <- nycomplaints_clean %>% 
  group_by(zip) %>% 
  tally()

types_of_complaints_by_zip <- nycomplaints_clean %>% 
  group_by(zip, complaint_type) %>% 
  tally()

most_complaints <- nycomplaints_spatial %>%
  group_by(geometry, zip, complaint_type) %>% 
  tally() %>% 
  slice_max(n) %>% 
  filter(!st_is_empty(geometry) & n > 1)

#showing the aggregate types of complaints that people file (over 1000 total)
types_of_complaints_by_zip %>% 
  group_by(complaint_type) %>% 
  summarise(count = sum(n)) %>% 
  filter(count > 1000) %>% 
  ggplot(aes(y = complaint_type, x = count)) +
  geom_col()

#most complaints map
most_complaints_map <- nycomplaints_spatial %>%
  group_by(geometry, zip, complaint_type) %>% 
  tally() %>% 
  slice_max(n) %>% 
  filter(!st_is_empty(geometry) & n > 1) %>% #filtering out only rows with geometry and complaints more than 1
  ggplot(aes(geometry = geometry, fill = complaint_type)) +
  geom_sf(color = NA) +
  theme_void()
print(most_complaints_map)

#by population percents map
pop_percent <- nyzip_census %>% 
  filter(variable != "med_hhinc") %>% 
  mutate(percent = 100 * (estimate / population_est)) %>% 
  ggplot(aes(fill = percent)) +
  facet_wrap(~variable) +
  geom_sf(color = NA) +
  theme_void()
print(pop_percent)

#pulling out median income by zip
income_by_zip <- nycomplaints_spatial %>% 
  select(zip, med_hhinc) %>% 
  group_by(zip, med_hhinc) %>% 
  summarise()


#median income map
income_map <- nyzip_census_wide %>%
  ggplot(aes(fill = med_hhinc)) +
  geom_sf(color = NA) +
  theme_void() +
  scale_fill_viridis_c(trans = "sqrt")
print(income_map)
```

## 4. Try to make one properly-polished plot or map

-   

    a.  *Either* choose one of the plots you've already made *or* create a new plot or map using the data. Work it up to something more finished. Aim for something that is both properly informative and as polished as you can make it.

-   

    b.  Try saving this plot or map using `ggsave()`. Create it as a PNG or PDF file. Make its dimensions (width and height) the right size for what you draw.
    
```{r}
complaints_income <- most_complaints %>% 
  left_join(income_by_zip, by = "zip")

income_summary <- summary(complaints_income$med_hhinc) %>% 
  unname()
income_summary <- income_summary[-4] #removing mean value
income_summary[1] <- income_summary[1] - 1 #ensuring that the minimum value is included
  
complaints_income$income_summary <- cut(complaints_income$med_hhinc, breaks = income_summary, 
    labels = c("21,447 - 54,184", "54,184 - 69,385", "69,385 - 92,840", "Above 92,840"))

income_nice <- complaints_income %>% 
  ggplot(aes(fill = income_summary, geometry = geometry)) +
  geom_sf() +
  scale_fill_brewer(palette = 16) +
  theme_void() +
  labs(title = "NYC Zip Codes by Income Quartiles",
       fill = "Income Quartiles") +
  theme(legend.position = "bottom", plot.title = element_text(size = 15, hjust = 0.5))

complaint_type_nice <- complaints_income %>% 
  ggplot(aes(fill = complaint_type, geometry = geometry)) +
  geom_sf() +
  scale_fill_viridis_d() +
  theme_void() +
  labs(title = "NYC Zip Codes by Main Complaint Type",
       fill = "Complaint Types") +
  theme(legend.position = "bottom", plot.title = element_text(size = 15, hjust = 0.5))


complaint_census_map <- income_nice + complaint_type_nice + plot_annotation(title = "NYC Comparison of Incomes by Complaints",
                                                                            theme = theme(plot.title = element_text(size = 20, hjust = 0.5)))
ggsave(here("complaint_census_map.pdf"), plot = complaint_census_map, device = "pdf", width = 15, height = 8)
```

Attempted to use ggpattern to overlay patterns based on complaint type but my computer either broke each time or wouldn't parse out 7 types of patterns

```{r}
complaints_income %>% 
  ggplot(aes(geometry = geometry)) +
  geom_sf_pattern(aes(pattern = complaint_type, fill = income_summary)
  ) +
  scale_pattern_manual(
    values = c(
      "Cultural Affairs" = "pch",
      "Housing and Buildings" = "circle",
      "Transportation" = "crosshatch",
      "Utilities" = "stripe",
      "Finance" = "none",
      "Quality of Life" = "rose",
      "Immigration" = "regular_polygon")) +
  theme_void()
```

