---
title: "Problem Set 12: NYC Complaints Data Analysis"
author: "<Judah Altman>"
date: "`r Sys.Date()`"
format: html
---

```{r}
#| label = "load-libraries-data",
#| warning = FALSE,
#| message = FALSE
library(tidyverse)
library(here)
library(socviz)

## Mapping
library(sf)

## Census
## Make sure you have a census API key! You only need one.
## https://api.census.gov/data/key_signup.html
library(tidycensus)
options(tigris_use_cache = TRUE)

## Install the data package:
## remotes::install_github("kjhealy/nycomplaints")
library(nycomplaints)
```

# The dataset

The dataset is documented [on its website](https://kjhealy.github.io/nycomplaints/). The source data is from [NYC Open Data](https://data.cityofnewyork.us/City-Government/NYC-Council-Constituent-Services/b9km-gdpy/about_data).

Read the documentation before working with the data.

# Explore the data

## The complaints data:

```{r}
nycomplaints
```

## Two useful tables

```{r}
nyc_zips
```

```{r}
census_vars
```

# Things to do

## 1. Briefly Describe the Dataset

-   Summarize the `nycomplaints` tibble in a way you think is informative to you.

```{r}
head(nycomplaints)
ncol(nycomplaints)
colnames(nycomplaints)
nrow(nycomplaints)
length(unique(nycomplaints$unique_key))
```

This is a tibble of constituent complaints registered by NYC Council district offices. It contains 341,299 rows and 11 columns, where each row (for the most part, to be investigated below) is a complaint. By columns, the unique key column identifies each complaint, the account refers to the council district that the complaint was filed in, the open and close date refer to the respective dates that the complaint was filed and when the case was closed, the complaint type categorizes the type of complaint, the descriptor gives a specific category within the broader type, and the remaining variables give the zip, borough, city, council district and community board of the constituent (note that the council district listed here might be different from the council district listed under account, as the account district refers to which district the complaint was filed in, while the council district refers to the council district that the constituent that filed the complaint lives in).

## 2 Look at each column/variable in a little more detail

-   Explore the basic structure of the data, by e.g. writing code to summarize the variables or investigate aspects of them that seem interesting. You can also create some initial rough exploratory plots if you like, in addition to tables or other summaries. As you go, explain here what it is you are doing and why.

```{r}
#overall data structure - checking the number of rows and columns and the types of data in the tibble
head(nycomplaints)
ncol(nycomplaints)
colnames(nycomplaints)
nrow(nycomplaints)

nycomplaints %>% 
  map(class)

#unique ID - checking for duplicate rows
length(unique(nycomplaints$unique_key))
nrow(nycomplaints)
duplicate_keys <- nycomplaints %>% 
  filter(duplicated(nycomplaints$unique_key)) %>% 
  select(unique_key) %>% 
  pull() %>% 
  str_c(collapse = "")

#checking if they are actually duplicates (they are)
duplicate_nycomplaints <- nycomplaints %>% 
  filter(str_detect(duplicate_keys, unique_key))
View(duplicate_nycomplaints)

#keeping only distinct ID keys (removing duplicates)
nycomplaints <- nycomplaints %>% 
  distinct(unique_key, .keep_all = TRUE)
length(nycomplaints)

#
```

## 3. Make some plots or maps

-   Following on from the tabular summarize, make some plots that you think show something interesting about the data. Make sure they look presentable and effective to you. Make a note about why you're doing each one. If you like, merge in data from elsewhere (e.g. from Census variables).

```{r}
nyzip_census <- readRDS(here("nyzip_census.rds"))
```



## 4. Try to make one properly-polished plot or map

-   

    a.  *Either* choose one of the plots you've already made *or* create a new plot or map using the data. Work it up to something more finished. Aim for something that is both properly informative and as polished as you can make it.

-   

    b.  Try saving this plot or map using `ggsave()`. Create it as a PNG or PDF file. Make its dimensions (width and height) the right size for what you draw.
